---
title: "REPRORODUCIBLE RESEARCH WITH BITCOIN FORECASTING"
author: "AYOUB"
date: "`r Sys.Date()`"
output: pdf_document
---

## BITCOIN TIME SERIES


Bitcoin is the longest running and most well known cryptocurrency, first released as open source in 2009 by the anonymous Satoshi Nakamoto. Bitcoin serves as a decentralized medium of digital exchange, with transactions verified and recorded in a public distributed ledger (the blockchain) without the need for a trusted record keeping authority or central intermediary. Transaction blocks contain a SHA-256 cryptographic hash of previous transaction blocks, and are thus "chained" together, serving as an immutable record of all transactions that have ever occurred. As with any currency/commodity on the market, bitcoin trading and financial instruments soon followed public adoption of bitcoin and continue to grow.

*I will also diivide my document into two parts*

-   Data exploration. -Time series analysis.

*Taking my time series further i will use*

-ARIMA -SARIMA -PROPHET -To further the study my data is upto date to 2023 august.

*THE NEWLY ADDED MODELS* From my analysis am trying to improve the auhors work by adding the following two new models. 1. SARIMA 2. WAVENET MODEL


## LIBRARIES USED IN THIS ANALYSIS

1.  A character vector named lib is created to store the names of the desired R packages. Each package name is enclosed in double quotation marks and separated by commas.

2.  The lapply() function is utilized to iterate over the elements of the lib vector. For each package name in the vector, the library() function is applied.

3.  This loads the respective package into the R environment. The character.only argument is set to TRUE to indicate that only character strings (package names) are being passed. The purpose of running this code is to ensure that the listed packages are loaded and available for use within the R script. It's important to note the following:

Some packages, such as rnn, TSrepr, sarima, tensorflow, keras, aTSA, xgboost, tidymodels, timetk, fpp3, and biwavelet, may require additional installation steps, dependencies, or prerequisites to be set up properly. An active internet connection is necessary, as the packages are downloaded and installed from the CRAN repository if they're not already present.

```{r echo=FALSE}
lib <- c("readr", "tidyverse", "tidyr", "rnn", "dplyr", "ggplot2", "plotly", "gganimate", "TSrepr", "sarima", "tensorflow", "keras","forecast","aTSA","xgboost","tidymodels","timetk","fpp3","biwavelet")
#lapply(lib, install.packages, character.only = TRUE) installation of packages if you lack them
lapply(lib, library, character.only = TRUE)

```

## DATA IMPORTATION

2.  <div>

    1.  This code reads a CSV file named "btc_history.csv" containing historical data for Bitcoin (BTC) and performs some data manipulation. Let's break down each line of code:
    2.  **`btc_history <- read_csv("btc_history.csv")`**:
        -   This line uses the **`read_csv()`** function from the "readr" package (assumed to be loaded in your environment) to read the contents of a CSV file named "btc_history.csv" into a data frame named **`btc_history`**.

        -   The **`read_csv()`** function is commonly used for reading comma-separated values (CSV) files while maintaining the data's structure.
    3.  **`btc_history$Date <- as.Date(btc_history$Date, "%b %d, %Y")`**:
        -   This line modifies the **`Date`** column in the **`btc_history`** data frame by converting it to a date format.

        -   **`btc_history$Date`** refers to the **`Date`** column in the data frame.

        -   The **`as.Date()`** function is used to convert the dates in the specified format ("%b %d, %Y") to the standard Date class in R. The format string corresponds to the format used in the column.

            -   **`%b`** represents abbreviated month names (e.g., Jan, Feb, \...).

            -   **`%d`** represents day of the month as a zero-padded decimal number (01, 02, \..., 31).

            -   **`%Y`** represents the year with century as a decimal number (e.g., 2023).

        -   The result of the **`as.Date()`** conversion is then assigned back to the **`Date`** column in the **`btc_history`** data frame.

    </div>

```{r}
btc_history <- read_csv("btc_history.csv")
btc_history$Date <- as.Date(btc_history$Date, "%b %d, %Y")
```
